A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
fix

\input{tag/counts.tex}
\input{tag/corpus.tex}
\input{tag/side-effects.tex}
\input{tag/operations.tex}
\input{tag/events.tex}
\input{tag/analysisStable.tex}
\input{tag/enrichDataset.tex}
\input{tag/patterns.tex}

@@

\newcommand{\eval}{\c{eval}}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}

\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

R supports reflection and meta-programming. The function \c{substitute(e,envir)}
yields the parse tree of the expression \c{e} after performing substitutions
defined by the bindings in \c{envir}.

@@

\noindent
R allows programmatic manipulation of parse trees, which are themselves
first class objects. They are evaluated using the \c{eval()}
function. There are three alternative forms  of \eval in the \c{base} package:
\c{evalq}, \c{eval.parent}, \c{local}.

\begin{lstlisting}
 eval(expr,
      envir = parent.frame(),
      enclos = if(is.list(envir) || is.pairlist(envir))
               parent.frame() else baseenv())
\end{lstlisting}

Arguments are as follows: \c{expr} is the value to be evaluated, \c{envir}
is the environment in which it is to be evaluated, and \c{enclos} is
relevant when \c{envir} is a list or a data frame, then it specifies where R
looks for objects not found in \c{envir}. This can be \c{NULL} (interpreted
as the base package environment, \c{baseenv()}) or an environment.

When called, \eval evaluates \c{expr} in the environment specified by
\c{envir} and return the computed value. If \c{envir} is missing, then the
default is \c{parent.frame()} (the environment where the call to \eval was
made).  Values to be evaluated can be of types \c{call} (AST for function
calls) or \c{expression} (AST nodes for arbitrary expressions), \c{name}
(when the symbol is looked up in the current scope and its binding is
evaluated), promise, or any of the basic types such as vectors and
environments (which are returned unchanged).

@@

@@

\begin{figure*}[!tb]\centering\includegraphics[width=.8\linewidth]
  {corpus.pdf}\caption{CRAN packages}\label{fig:corpus}
\end{figure*}

@@

@@

There are three sources of runnable code in a R package. Next to the package
\emph{tests},\footnote{We had to exclude most tests because the {testthat}
  harness uses \eval and thus causes the entire test to register as an \eval
  call.} there are also \emph{examples} and \emph{vignettes}. Examples are R
code snippets extracted from package documentation into scripts files.
Vignettes are long-form description of package functionality written using
literate-programming style in \LaTeX\xspace or Rmarkdown with chunks of R
code that can be similarly extracted into R scripts. R provides builtin
functions for tangling the embedded code into files.  The selected packages
contain \CorpusPackagePrograms programs with \CorpusPackageProgramsCodeRnd
lines of R code: \CorpusExamplesProgramsRnd examples with
\summaryrnd{CorpusExamplesCode}\footnote{This notation represents $sum$
  ($min$ / $mean$ / $max$)} lines of code, and \CorpusVignettesProgramsRnd
vignettes.%%% with \summaryrnd{CorpusVignettesCode} lines of code.

@@

To represent end-user code in the corpus, we turned to
Kaggle,\footnote{\url{https://www.kaggle.com}} an online platform for
data-science and machine-learning. The website allows people to share
data-science competitions, a data-analysis problem together with data, for
which users try to find the best solution. The solutions, called {kernels},
are then shared back to the platform in the form of either plain scripts or
as notebooks.  One of the most popular competition is about predicting
passenger survival on
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}} with \CorpusKaggle
kernels in R (over 1/4 of all available R kernels) which we used for our
corpus.

>>>>

%\input{tag/counts.tex}
\newcommand{\CoreEvalCallCountRnd}{6.3M\xspace}
\newcommand{\CoreEvalCallCount}{6,282,429\xspace}
\newcommand{\CoreEvalSiteCountRnd}{185\xspace}
\newcommand{\CoreEvalSiteCount}{185\xspace}
\newcommand{\CoreEvalParentCallCountRnd}{186K\xspace}
\newcommand{\CoreEvalParentCallCount}{186,275\xspace}
\newcommand{\CoreEvalParentSiteCountRnd}{13\xspace}
\newcommand{\CoreEvalParentSiteCount}{13\xspace}
\newcommand{\CoreEvalqCallCountRnd}{132K\xspace}
\newcommand{\CoreEvalqCallCount}{132,891\xspace}
\newcommand{\CoreEvalqSiteCountRnd}{20K\xspace}
\newcommand{\CoreEvalqSiteCount}{20,460\xspace}
\newcommand{\CoreLocalCallCountRnd}{175K\xspace}
\newcommand{\CoreLocalCallCount}{175,079\xspace}
\newcommand{\CoreLocalSiteCountRnd}{920\xspace}
\newcommand{\CoreLocalSiteCount}{920\xspace}
\newcommand{\PackageEvalCallCountRnd}{1.6M\xspace}
\newcommand{\PackageEvalCallCount}{1,582,738\xspace}
\newcommand{\PackageEvalSiteCountRnd}{1.5K\xspace}
\newcommand{\PackageEvalSiteCount}{1,459\xspace}
\newcommand{\PackageEvalParentCallCountRnd}{13K\xspace}
\newcommand{\PackageEvalParentCallCount}{13,230\xspace}
\newcommand{\PackageEvalParentSiteCountRnd}{249\xspace}
\newcommand{\PackageEvalParentSiteCount}{249\xspace}
\newcommand{\PackageEvalqCallCountRnd}{2\xspace}
\newcommand{\PackageEvalqCallCount}{2\xspace}
\newcommand{\PackageEvalqSiteCountRnd}{1\xspace}
\newcommand{\PackageEvalqSiteCount}{1\xspace}
\newcommand{\PackageLocalCallCountRnd}{2.2K\xspace}
\newcommand{\PackageLocalCallCount}{2,228\xspace}
\newcommand{\PackageLocalSiteCountRnd}{15\xspace}
\newcommand{\PackageLocalSiteCount}{15\xspace}
\newcommand{\KaggleEvalCallCountRnd}{0\xspace}
\newcommand{\KaggleEvalCallCount}{0\xspace}
\newcommand{\KaggleEvalSiteCountRnd}{0\xspace}
\newcommand{\KaggleEvalSiteCount}{0\xspace}
\newcommand{\KaggleEvalParentCallCountRnd}{0\xspace}
\newcommand{\KaggleEvalParentCallCount}{0\xspace}
\newcommand{\KaggleEvalParentSiteCountRnd}{0\xspace}
\newcommand{\KaggleEvalParentSiteCount}{0\xspace}
\newcommand{\KaggleEvalqCallCountRnd}{0\xspace}
\newcommand{\KaggleEvalqCallCount}{0\xspace}
\newcommand{\KaggleEvalqSiteCountRnd}{0\xspace}
\newcommand{\KaggleEvalqSiteCount}{0\xspace}
\newcommand{\KaggleLocalCallCountRnd}{0\xspace}
\newcommand{\KaggleLocalCallCount}{0\xspace}
\newcommand{\KaggleLocalSiteCountRnd}{0\xspace}
\newcommand{\KaggleLocalSiteCount}{0\xspace}
\newcommand{\CoreAllCallCountRnd}{6.8M\xspace}
\newcommand{\CoreAllCallCount}{6,776,674\xspace}
\newcommand{\CoreAllSiteCountRnd}{21.6K\xspace}
\newcommand{\CoreAllSiteCount}{21,578\xspace}
\newcommand{\KaggleAllCallCountRnd}{0\xspace}
\newcommand{\KaggleAllCallCount}{0\xspace}
\newcommand{\KaggleAllSiteCountRnd}{0\xspace}
\newcommand{\KaggleAllSiteCount}{0\xspace}
\newcommand{\PackageAllCallCountRnd}{1.6M\xspace}
\newcommand{\PackageAllCallCount}{1,598,198\xspace}
\newcommand{\PackageAllSiteCountRnd}{1.7K\xspace}
\newcommand{\PackageAllSiteCount}{1,724\xspace}
\newcommand{\AllEvalCallCountRnd}{7.9M\xspace}
\newcommand{\AllEvalCallCount}{7,865,167\xspace}
\newcommand{\AllEvalSiteCountRnd}{1.6K\xspace}
\newcommand{\AllEvalSiteCount}{1,644\xspace}
\newcommand{\AllEvalParentCallCountRnd}{199.5K\xspace}
\newcommand{\AllEvalParentCallCount}{199,505\xspace}
\newcommand{\AllEvalParentSiteCountRnd}{262\xspace}
\newcommand{\AllEvalParentSiteCount}{262\xspace}
\newcommand{\AllEvalqCallCountRnd}{132.9K\xspace}
\newcommand{\AllEvalqCallCount}{132,893\xspace}
\newcommand{\AllEvalqSiteCountRnd}{20.5K\xspace}
\newcommand{\AllEvalqSiteCount}{20,461\xspace}
\newcommand{\AllLocalCallCountRnd}{177.3K\xspace}
\newcommand{\AllLocalCallCount}{177,307\xspace}
\newcommand{\AllLocalSiteCountRnd}{935\xspace}
\newcommand{\AllLocalSiteCount}{935\xspace}
\newcommand{\AllAllCallCountRnd}{8.4M\xspace}
\newcommand{\AllAllCallCount}{8,374,872\xspace}
\newcommand{\AllAllSiteCountRnd}{23.3K\xspace}
\newcommand{\AllAllSiteCount}{23,302\xspace}
\newcommand{\TotalFileCountRnd}{18.3K\xspace}
\newcommand{\TotalFileCount}{18,258\xspace}
\newcommand{\NoEvalFileCountRnd}{89\xspace}
\newcommand{\NoEvalFileCount}{89\xspace}
\newcommand{\NoEvalFilePerc}{0.5\%\xspace}
\newcommand{\CoreEvalFileCountRnd}{12.3K\xspace}
\newcommand{\CoreEvalFileCount}{12,307\xspace}
\newcommand{\CoreEvalFilePerc}{67.4\%\xspace}
\newcommand{\PackageEvalFileCountRnd}{0\xspace}
\newcommand{\PackageEvalFileCount}{0\xspace}
\newcommand{\PackageEvalFilePerc}{0\%\xspace}
\newcommand{\AllEvalFileCountRnd}{5.9K\xspace}
\newcommand{\AllEvalFileCount}{5,862\xspace}
\newcommand{\AllEvalFilePerc}{32.1\%\xspace}
\newcommand{\EightyCoreEvalFileCountRnd}{3.8K\xspace}
\newcommand{\EightyCoreEvalFileCount}{3,769\xspace}
\newcommand{\EightyCoreEvalFilePerc}{64.3\%\xspace}
\newcommand{\TopTenPackageCallCountRnd}{1.4M\xspace}
\newcommand{\TopTenPackageCallCount}{1,425,902\xspace}
\newcommand{\TopTenPackageCallPerc}{89.2\%\xspace}
\newcommand{\TopTenPackageSiteCountRnd}{282\xspace}
\newcommand{\TopTenPackageSiteCount}{282\xspace}
\newcommand{\TopTenPackageSitePerc}{16\%\xspace}
\newcommand{\TopTenPackageNameA}{ggplot2\xspace}
\newcommand{\TopTenPackageCallsiteCountARnd}{2\xspace}
\newcommand{\TopTenPackageCallsiteCountA}{2\xspace}
\newcommand{\TopTenPackageCallCountARnd}{987K\xspace}
\newcommand{\TopTenPackageCallCountA}{987,946\xspace}
\newcommand{\TopTenPackageCallPercA}{61\%\xspace}
\newcommand{\TopTenPackageNameB}{magrittr\xspace}
\newcommand{\TopTenPackageCallsiteCountBRnd}{5\xspace}
\newcommand{\TopTenPackageCallsiteCountB}{5\xspace}
\newcommand{\TopTenPackageCallCountBRnd}{139K\xspace}
\newcommand{\TopTenPackageCallCountB}{139,346\xspace}
\newcommand{\TopTenPackageCallPercB}{8.7\%\xspace}
\newcommand{\TopTenPackageNameC}{plyr\xspace}
\newcommand{\TopTenPackageCallsiteCountCRnd}{11\xspace}
\newcommand{\TopTenPackageCallsiteCountC}{11\xspace}
\newcommand{\TopTenPackageCallCountCRnd}{68K\xspace}
\newcommand{\TopTenPackageCallCountC}{68,671\xspace}
\newcommand{\TopTenPackageCallPercC}{4.3\%\xspace}
\newcommand{\TopTenPackageNameD}{glue\xspace}
\newcommand{\TopTenPackageCallsiteCountDRnd}{5\xspace}
\newcommand{\TopTenPackageCallsiteCountD}{5\xspace}
\newcommand{\TopTenPackageCallCountDRnd}{62K\xspace}
\newcommand{\TopTenPackageCallCountD}{62,564\xspace}
\newcommand{\TopTenPackageCallPercD}{3.9\%\xspace}
\newcommand{\TopTenPackageNameE}{data.table\xspace}
\newcommand{\TopTenPackageCallsiteCountERnd}{31\xspace}
\newcommand{\TopTenPackageCallsiteCountE}{31\xspace}
\newcommand{\TopTenPackageCallCountERnd}{47K\xspace}
\newcommand{\TopTenPackageCallCountE}{47,464\xspace}
\newcommand{\TopTenPackageCallPercE}{3\%\xspace}
\newcommand{\TopTenPackageNameF}{statnet.common\xspace}
\newcommand{\TopTenPackageCallsiteCountFRnd}{12\xspace}
\newcommand{\TopTenPackageCallsiteCountF}{12\xspace}
\newcommand{\TopTenPackageCallCountFRnd}{30K\xspace}
\newcommand{\TopTenPackageCallCountF}{30,581\xspace}
\newcommand{\TopTenPackageCallPercF}{1.9\%\xspace}
\newcommand{\TopTenPackageNameG}{VGAM\xspace}
\newcommand{\TopTenPackageCallsiteCountGRnd}{97\xspace}
\newcommand{\TopTenPackageCallsiteCountG}{97\xspace}
\newcommand{\TopTenPackageCallCountGRnd}{23K\xspace}
\newcommand{\TopTenPackageCallCountG}{23,796\xspace}
\newcommand{\TopTenPackageCallPercG}{1.5\%\xspace}
\newcommand{\TopTenPackageNameH}{copula\xspace}
\newcommand{\TopTenPackageCallsiteCountHRnd}{21\xspace}
\newcommand{\TopTenPackageCallsiteCountH}{21\xspace}
\newcommand{\TopTenPackageCallCountHRnd}{22K\xspace}
\newcommand{\TopTenPackageCallCountH}{22,794\xspace}
\newcommand{\TopTenPackageCallPercH}{1.4\%\xspace}
\newcommand{\TopTenPackageNameI}{nlme\xspace}
\newcommand{\TopTenPackageCallsiteCountIRnd}{97\xspace}
\newcommand{\TopTenPackageCallsiteCountI}{97\xspace}
\newcommand{\TopTenPackageCallCountIRnd}{21K\xspace}
\newcommand{\TopTenPackageCallCountI}{21,483\xspace}
\newcommand{\TopTenPackageCallPercI}{1.3\%\xspace}
\newcommand{\TopTenPackageNameJ}{R6\xspace}
\newcommand{\TopTenPackageCallsiteCountJRnd}{1\xspace}
\newcommand{\TopTenPackageCallsiteCountJ}{1\xspace}
\newcommand{\TopTenPackageCallCountJRnd}{21K\xspace}
\newcommand{\TopTenPackageCallCountJ}{21,257\xspace}
\newcommand{\TopTenPackageCallPercJ}{1.3\%\xspace}

\input{tag/corpus.tex}
%\input{tag/side-effects.tex}
\newcommand{\AllWritesRnd}{9.4G\xspace}
\newcommand{\AllWrites}{9,427,046,091\xspace}
\newcommand{\EvalWritesRnd}{1.1G\xspace}
\newcommand{\EvalWrites}{1,065,212,013\xspace}
\newcommand{\EvalSideEffectingWritesRnd}{45M\xspace}
\newcommand{\EvalSideEffectingWrites}{45,701,209\xspace}
\newcommand{\EvalSideEffectingWritesEvalPerc}{4.3\%\xspace}
\newcommand{\EvalSideEffectingWritesAllPerc}{.5\%\xspace}
\newcommand{\SideEffectingCoreCallPerc}{4.4\%\xspace}
\newcommand{\SideEffectingPackageCallPerc}{7.5\%\xspace}

%\input{tag/operations.tex}
\newcommand{\AllValueInputEvalCallPerc}{17\%\xspace}
\newcommand{\AllExpressionInputEvalCallPerc}{82\%\xspace}
\newcommand{\AllExpressionInputEvalSitePerc}{14\%\xspace}
\newcommand{\AllValueInputEvalSitePerc}{3\%\xspace}
\newcommand{\AllPolymorphicInputEvalSitePerc}{82\%\xspace}
\newcommand{\PackageEnvironmentInputEvalCallPerc}{62\%\xspace}
\newcommand{\AllClosureCallCountRnd}{2.9G\xspace}
\newcommand{\AllClosureCallCount}{2,886,992,072\xspace}
\newcommand{\EvalClosureCallCountRnd}{125M\xspace}
\newcommand{\EvalClosureCallCount}{125,469,701\xspace}
\newcommand{\AllNativeCallCountRnd}{463M\xspace}
\newcommand{\AllNativeCallCount}{463,803,879\xspace}
\newcommand{\EvalNativeCallCountRnd}{31M\xspace}
\newcommand{\EvalNativeCallCount}{31,164,219\xspace}

%\input{tag/events.tex}
\newcommand{\AllEventCountRnd}{39G\xspace}
\newcommand{\AllEventCount}{38,991,360,760\xspace}
\newcommand{\EvalEventCountRnd}{2G\xspace}
\newcommand{\EvalEventCount}{1,973,736,718\xspace}
\newcommand{\EventsMaxCountRnd}{17M\xspace}
\newcommand{\EventsMaxCount}{17,656,694\xspace}
\newcommand{\EvalEventAllPerc}{5\%\xspace}
\newcommand{\EventsMinRangeARnd}{0\xspace}
\newcommand{\EventsMinRangeA}{0\xspace}
\newcommand{\EventsMaxRangeARnd}{100\xspace}
\newcommand{\EventsMaxRangeA}{100\xspace}
\newcommand{\EventsCoreEvalCountARnd}{5M\xspace}
\newcommand{\EventsCoreEvalCountA}{5,659,053\xspace}
\newcommand{\EventsCoreEvalPercA}{88\%\xspace}
\newcommand{\EventsPackageEvalCountARnd}{1.5M\xspace}
\newcommand{\EventsPackageEvalCountA}{1,514,222\xspace}
\newcommand{\EventsPackageEvalPercA}{93\%\xspace}
\newcommand{\EventsMinRangeBRnd}{101\xspace}
\newcommand{\EventsMinRangeB}{101\xspace}
\newcommand{\EventsMaxRangeBRnd}{10K\xspace}
\newcommand{\EventsMaxRangeB}{10,000\xspace}
\newcommand{\EventsCoreEvalCountBRnd}{769K\xspace}
\newcommand{\EventsCoreEvalCountB}{769,341\xspace}
\newcommand{\EventsCoreEvalPercB}{12\%\xspace}
\newcommand{\EventsPackageEvalCountBRnd}{100K\xspace}
\newcommand{\EventsPackageEvalCountB}{100,141\xspace}
\newcommand{\EventsPackageEvalPercB}{6\%\xspace}
\newcommand{\EventsMinRangeCRnd}{10K\xspace}
\newcommand{\EventsMinRangeC}{10,001\xspace}
\newcommand{\EventsMaxRangeCRnd}{20M\xspace}
\newcommand{\EventsMaxRangeC}{20,000,000\xspace}
\newcommand{\EventsCoreEvalCountCRnd}{3K\xspace}
\newcommand{\EventsCoreEvalCountC}{2,997\xspace}
\newcommand{\EventsCoreEvalPercC}{0\%\xspace}
\newcommand{\EventsPackageEvalCountCRnd}{14K\xspace}
\newcommand{\EventsPackageEvalCountC}{14,209\xspace}
\newcommand{\EventsPackageEvalPercC}{.9\%\xspace}

%\input{tag/analysisStable.tex}
\newcommand{\UniqueCallsRnd}{1.3M\xspace}
\newcommand{\UniqueCalls}{1,277,948\xspace}
\newcommand{\MatchArgNumberRnd}{5M\xspace}
\newcommand{\MatchArgNumber}{5,029,351\xspace}
\newcommand{\MatchArgPercent}{60\%\xspace}
\newcommand{\NbAstOnePercent}{15\%\xspace}
\newcommand{\NbAstOneCallSitePercent}{96\%\xspace}
\newcommand{\AstSizeNineFiveRnd}{42\xspace}
\newcommand{\AstSizeNineFive}{42\xspace}
\newcommand{\AstSizeNineFiveExactPercent}{95\%\xspace}
\newcommand{\DefaultEnvirExprRnd}{1.1M\xspace}
\newcommand{\DefaultEnvirExpr}{1,054,124\xspace}
\newcommand{\DefaultEnvirExprPercent}{12\%\xspace}
\newcommand{\DefaultEnvirCallSitesPercent}{9.3\%\xspace}
\newcommand{\DefaultEnvirCoreExprPercent}{10\%\xspace}
\newcommand{\DefaultEnvirPackagesExprPercent}{20\%\xspace}
\newcommand{\CoreEnvClassA}{2\xspace}
\newcommand{\CoreCountARnd}{5M\xspace}
\newcommand{\CoreCountA}{5,079,433\xspace}
\newcommand{\CoreProportionA}{83\%\xspace}
\newcommand{\CoreEnvClassB}{list\xspace}
\newcommand{\CoreCountBRnd}{229K\xspace}
\newcommand{\CoreCountB}{229,249\xspace}
\newcommand{\CoreProportionB}{3.8\%\xspace}
\newcommand{\CoreEnvClassC}{3\xspace}
\newcommand{\CoreCountCRnd}{199K\xspace}
\newcommand{\CoreCountC}{199,497\xspace}
\newcommand{\CoreProportionC}{3.3\%\xspace}
\newcommand{\CoreEnvClassD}{global+\xspace}
\newcommand{\CoreCountDRnd}{183K\xspace}
\newcommand{\CoreCountD}{183,631\xspace}
\newcommand{\CoreProportionD}{3\%\xspace}
\newcommand{\CoreEnvClassE}{1+\xspace}
\newcommand{\CoreCountERnd}{175K\xspace}
\newcommand{\CoreCountE}{175,151\xspace}
\newcommand{\CoreProportionE}{2.9\%\xspace}
\newcommand{\CoreEnvClassF}{global\xspace}
\newcommand{\CoreCountFRnd}{124K\xspace}
\newcommand{\CoreCountF}{124,673\xspace}
\newcommand{\CoreProportionF}{2.1\%\xspace}
\newcommand{\CoreEnvClassG}{sys.call(sys.parent())\xspace}
\newcommand{\CoreCountGRnd}{32K\xspace}
\newcommand{\CoreCountG}{32,349\xspace}
\newcommand{\CoreProportionG}{.5\%\xspace}
\newcommand{\CoreEnvClassH}{loop\xspace}
\newcommand{\CoreCountHRnd}{24K\xspace}
\newcommand{\CoreCountH}{24,286\xspace}
\newcommand{\CoreProportionH}{.4\%\xspace}
\newcommand{\PackagesEnvClassA}{global+\xspace}
\newcommand{\PackagesCountARnd}{943K\xspace}
\newcommand{\PackagesCountA}{943,682\xspace}
\newcommand{\PackagesProportionA}{74\%\xspace}
\newcommand{\PackagesEnvClassB}{3\xspace}
\newcommand{\PackagesCountBRnd}{71K\xspace}
\newcommand{\PackagesCountB}{71,037\xspace}
\newcommand{\PackagesProportionB}{5.6\%\xspace}
\newcommand{\PackagesEnvClassC}{2\xspace}
\newcommand{\PackagesCountCRnd}{67K\xspace}
\newcommand{\PackagesCountC}{67,555\xspace}
\newcommand{\PackagesProportionC}{5.3\%\xspace}
\newcommand{\PackagesEnvClassD}{5\xspace}
\newcommand{\PackagesCountDRnd}{49K\xspace}
\newcommand{\PackagesCountD}{49,782\xspace}
\newcommand{\PackagesProportionD}{3.9\%\xspace}
\newcommand{\PackagesEnvClassE}{list\xspace}
\newcommand{\PackagesCountERnd}{44K\xspace}
\newcommand{\PackagesCountE}{44,194\xspace}
\newcommand{\PackagesProportionE}{3.5\%\xspace}
\newcommand{\PackagesEnvClassF}{2+\xspace}
\newcommand{\PackagesCountFRnd}{32K\xspace}
\newcommand{\PackagesCountF}{32,931\xspace}
\newcommand{\PackagesProportionF}{2.6\%\xspace}
\newcommand{\PackagesEnvClassG}{global\xspace}
\newcommand{\PackagesCountGRnd}{22K\xspace}
\newcommand{\PackagesCountG}{22,825\xspace}
\newcommand{\PackagesProportionG}{1.8\%\xspace}
\newcommand{\PackagesEnvClassH}{base+\xspace}
\newcommand{\PackagesCountHRnd}{18K\xspace}
\newcommand{\PackagesCountH}{18,538\xspace}
\newcommand{\PackagesProportionH}{1.5\%\xspace}
\newcommand{\NbEnclosEnvirRnd}{274K\xspace}
\newcommand{\NbEnclosEnvir}{274,370\xspace}
\newcommand{\EnclosEnvirPercent}{3.3\%\xspace}
\newcommand{\NbSysCallEnvirRnd}{33K\xspace}
\newcommand{\NbSysCallEnvir}{33,650\xspace}
\newcommand{\SysCallEnvirPercent}{.4\%\xspace}
\newcommand{\NbMonomorphicRnd}{23.1K\xspace}
\newcommand{\NbMonomorphic}{23,069\xspace}
\newcommand{\PercentMonomorphic}{99\%\xspace}
\newcommand{\PercentParsedEvals}{.9\%\xspace}
\newcommand{\PercentParsedCallSites}{1\%\xspace}
\newcommand{\NbParseFilesRnd}{10\xspace}
\newcommand{\NbParseFiles}{10\xspace}
\newcommand{\PercentParseFileCallSites}{.01\%\xspace}
\newcommand{\PasteParsedEvals}{39\%\xspace}
\newcommand{\PasteParsedCallSite}{51\%\xspace}
\newcommand{\NbCoreEvalCallSitesRnd}{310\xspace}
\newcommand{\NbCoreEvalCallSites}{310\xspace}
\newcommand{\MaxCallSitePackageRnd}{2.4K\xspace}
\newcommand{\MaxCallSitePackage}{2,372\xspace}
\newcommand{\OneCallSitePackageRnd}{48\xspace}
\newcommand{\OneCallSitePackage}{48\xspace}
\newcommand{\NoCallSitePackagesRnd}{233\xspace}
\newcommand{\NoCallSitePackages}{233\xspace}

%\input{tag/enrichDataset.tex}
\newcommand{\UndefinedEvalsRnd}{285K\xspace}
\newcommand{\UndefinedEvals}{285,289\xspace}
\newcommand{\PercentUndefinedEval}{3.3\%\xspace}
\newcommand{\UndefinedEvalExpressionsRnd}{262\xspace}
\newcommand{\UndefinedEvalExpressions}{262\xspace}
\newcommand{\UndefinedOutOfCorpusRnd}{0\xspace}
\newcommand{\UndefinedOutOfCorpus}{0\xspace}
\newcommand{\UndefinedOutOfCoprusPercent}{0\%\xspace}
\newcommand{\MaxCallSitesEnrich}{114\xspace}
\newcommand{\OneCallSitesEnrich}{61\xspace}
\newcommand{\NoCallSitesEnrich}{297\xspace}
%\input{tag/patterns.tex}
\newcommand{\PatternLibRnd}{4K\xspace}
\newcommand{\PatternLib}{4,278\xspace}
\newcommand{\PatternLibRatio}{.05\%\xspace}
\newcommand{\PatternLibSourceRnd}{370\xspace}
\newcommand{\PatternLibSource}{370\xspace}
\newcommand{\PatternLibLocalRnd}{100\xspace}
\newcommand{\PatternLibLocal}{100\xspace}
\newcommand{\PatternReadRnd}{59K\xspace}
\newcommand{\PatternRead}{59,776\xspace}
\newcommand{\PatternReadRatio}{.7\%\xspace}
\newcommand{\PatternReadSlotAccessRatio}{88\%\xspace}
\newcommand{\PatternReadDatatableRatio}{27\%\xspace}
\newcommand{\PatternAssignRnd}{68K\xspace}
\newcommand{\PatternAssign}{68,754\xspace}
\newcommand{\PatternAssignRatio}{.9\%\xspace}
\newcommand{\PatternAssignArrowRatio}{18\%\xspace}
\newcommand{\PatternAssignArrowMagrittrRatio}{59\%\xspace}
\newcommand{\PatternAssignSuperarrowRatio}{81\%\xspace}
\newcommand{\PatternAssignSuperarrowPlyrRatio}{99\%\xspace}
\newcommand{\PatternManualPackagesRnd}{20\xspace}
\newcommand{\PatternManualPackages}{20\xspace}
\newcommand{\PatternManualCallsitesRnd}{161\xspace}
\newcommand{\PatternManualCallsites}{161\xspace}

@@

\newcommand{\eval}{\texttt{eval}}
\newcommand{\evals}{{\sf eval}s\xspace}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}
\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

R supports reflection and meta-programming. The function \c{substitute(e,envir)}
yields the parse tree of the expression \c{e} after performing substitutions
defined by the bindings in the \c{envir} parameter.

@@

\noindent
R allows programmatic manipulation of parse trees, which are themselves
first class objects. They are evaluated using the \c{eval()}
function. There are three alternative forms  of \eval in the \c{base} package:
\c{evalq}, \c{eval.parent} and \c{local}.

\eval function in R takes three arguments as shown below.
\begin{lstlisting}
 eval(expr,
      envir = parent.frame(),
      enclos = if(is.list(envir) || is.pairlist(envir))
               parent.frame() else baseenv())
\end{lstlisting}

\c{expr} is the value to be evaluated, \c{envir}
is the environment in which it is to be evaluated, and \c{enclos} is
relevant when \c{envir} is a list or a data frame, then it specifies where R
looks for objects not found in \c{envir}. This can be \c{NULL} (interpreted
as the base package environment, \c{baseenv()}) or an environment.

When called, \eval evaluates \c{expr} in the environment specified by
\c{envir} and returns the computed value. If \c{envir} is missing, then the
default is \c{parent.frame()} (the environment where the call to \eval was
made).  Values to be evaluated can be of types \c{call} (AST for function
calls) or \c{expression} (AST nodes for arbitrary expressions), \c{name}
(when the symbol is looked up in the current scope and its binding is
evaluated), promise, or any of the basic types such as vectors and
environments (which are returned unchanged).

@@

\medskip\emph{Discussion:~} There are two outstanding differences that set
the design of \eval in R apart from JavaScript. On the usability side, \eval
does not take a string, but an AST and an environment. This is slightly more
involved than just taking a string and may discourage some casual users.
The second substantial difference is the control that R gives over the
choice of environment in which the AST is evaluated in. Beside traditional
choices such as the top level or current lexical environment, R allows
programmers to choose a fresh, empty, environment or any environment on the
call stack. This is a significant increase in expressive power as functions
that are otherwise not aware of the use of \eval may be deeply impacted.

@@

@@

\begin{figure}[!h]\centering\includegraphics[width=\linewidth]
  {corpus.pdf}\caption{CRAN packages}\label{fig:corpus}
\end{figure}

@@

There are three sources of runnable code in a R package. Next to the package
\emph{tests},\footnote{We had to exclude most tests because the {testthat}
  harness uses \eval and thus causes the entire test to register as an \eval
  call.} there are also \emph{examples} and \emph{vignettes}. Examples are R
code snippets extracted from package documentation into scripts files.
Vignettes are long-form description of package functionality written using
literate-programming style in \LaTeX\xspace or Rmarkdown with chunks of R
code that can be similarly extracted into R scripts. R provides builtin
functions for tangling the embedded code into files.  The selected packages
contain \CorpusPackagePrograms programs with \CorpusPackageProgramsCodeRnd
lines of R code: \CorpusExamplesProgramsRnd examples with
\CorpusExamplesCodeRnd lines of code, and \CorpusVignettesProgramsRnd
vignettes.

@@

To represent end-user code in the corpus, we turned to Kaggle, an online
platform for data-science and machine-learning. The website allows people to
share data-science competitions, a data-analysis problem together with data,
for which users try to find the best solution. The solutions, called
{kernels}, are then shared back to the platform in the form of either plain
scripts or as notebooks.  One of the most popular competition is about
predicting passenger survival on
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}} with \CorpusKaggle
kernels in R (over 1/4 of all available R kernels) which we used for our
corpus.

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats,\footnote{We use {\sf rmarkdown} to convert from notebooks to R.} we
found that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next
to various runtime exceptions, common problems were parse errors and
misspelled package names.  The final set contains \CorpusFinishedKaggle
kernels with \CorpusFinishedKaggleCodeRnd lines of R code.
